Practical Optional Types for Clojure

### 2015-02-23 intro.tex
use hmaps and let-aliasing in intro example

\begin{figure}
\inputminted[firstline=5,lastline=10]{clojure}{code/demo/src/demo/parent.clj}
\caption{A simple Typed Clojure program}
\label{fig:ex1}
\end{figure}

Figure~\ref{fig:ex1} presents a simple program demonstrating several
aspects of our system, from simple type annotations to explicit
handling of Java's \texttt{null} in interoperation, as well as an
extended form of occurrence typing and Clojure's \emph{type hints},
which are central to Typed Clojure's approach to interoperability.

In the remainder of this paper, we describe Typed Clojure's central
innovations, including Java interoperability, multimethods, and
heterogenously-typed immutable maps, following an example-driven
presentation in Section~\ref{sec:overview}. We then incrementatlly
present a core calculus for Typed Clojure covering all of these
features together in Section~\ref{sec:formal} and prove type
soundness(section \ref{sec:sound}). We then discuss the full
implementation of Typed Clojure, dubbed \coretyped{}, which
extends the formal model in many ways, and the experience gained from
its use in Section~\ref{sec:exp}. Finally, we discuss related work and
conclude.

>>>>

\begin{figure}
\inputminted[firstline=5,lastline=8]{clojure}{code/demo/src/demo/parent2.clj}
\caption{A simple Typed Clojure program}
\label{fig:ex1}
\end{figure}

Figure~\ref{fig:ex1} presents a simple program demonstrating many
aspects of our system, from simple type annotations to explicit
handling of Java's \texttt{null} in interoperation, as well as an
extended form of occurrence typing and Clojure's \emph{type hints},
which are central to Typed Clojure's approach to interoperability.

The \clj{parent} function has the type 
$$
\clj{['{:file (U nil File)} -> (U nil Str)]}
$$
which means that it takes a hash table whose \clj{:file} key maps to either
\clj{nil} or a \clj{File}, and it produces either \clj{nil} or a
\clj{String}. The \clj{parent} function uses the \clj{:file} keyword
as an accessor to get the file, checks that it isn't \clj{nil}, and
then obtains the parent by making a Java method call.
%
The annotation \clj{^File} is a type hint, which instructs the Clojure
compiler (running prior to Typed Clojure typechecking) to statically
resolve method calls to the relevant object using the Java \clj{File}
class, rather than using reflective method calls at runtime.

In the remainder of this paper, we describe how Typed Clojure's
central innovations, including Java interoperability, multimethods,
and heterogenously-typed immutable maps, enable this example and many
others. We begin with an example-driven presentation of the main type
system features in Section~\ref{sec:overview}. We then incrementatlly
present a core calculus for Typed Clojure covering all of these
features together in Section~\ref{sec:formal} and prove type
soundness(section \ref{sec:sound}). We then discuss the full
implementation of Typed Clojure, dubbed \coretyped{}, which extends
the formal model in many ways, and the experience gained from its use
in Section~\ref{sec:exp}. Finally, we discuss related work and
conclude.
